\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-02 paper/main.tex
Updates in Results

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow patterns~\cite{Olstad2019CiliaryDevelopment}.
To assess how these variations influence solute distribution,
we consider four alternative ventricular geometries:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a, b}).
In these four modified geometries, the total volume is reduced by
3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.

Constricting the connection between the anterior and middle ventricle reduces
solute transport to the anterior ventricle (ROI 5),
and the mean concentration in ROIs 2--4 and 6 slightly increases, compared to the original geometry
(\Crefpanels{fig:fig6}{c}). This is notwithstanding a
9.5\% reduction in ventral cilia force magnitude
(\Crefpanels{fig:fig6}{d}), which are the cilia that propel CSF and solute towards ROI 6.
Constricting the mid-hind brain connection
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5).
Faster transport to the anterior ventricle (ROI 5) and slower transport to the
posterior ventricle (ROI 6) is also the case for the geometry with a
shrunk middle ventricle.
This reduction in the transport to the posterior ventricle (ROI 6)
happens in spite of a 6.6\% increase in the ventral cilia forces.
Additionally, for the shrunk middle configuration, the transport upstream of CSF flow towards the
ventral-posterior region of the middle ventricle (ROI 3) is slowed down.

For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have reduced
mean concentrations. To summarize, our results show that the
ventricular geometry and the cross-sectional area of the ventricular ducts matter
for solute transport, with notable differences in transport resulting from
relatively modest geometry modifications. This may explain why
we observed larger differences in the cilia-deficient computational
model results compared to the experimental data in \Cref{fig:fig4},
as the model and the fish have slightly different geometries.
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}

>>>>

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow
patterns~\cite{Olstad2019CiliaryDevelopment}.  To assess how these
variations influence solute distribution, we consider four alternative
ventricular geometries: (i) constriction of the anterior-middle
ventricular duct (66 \% reduction in cross-sectional area); (ii) constriction of the middle-posterior duct (33 \% reduction); (iii) middle ventricle shrunk in the lateral direction (43 \% reduction); and (iv) all three combined (\Crefpanels{fig:fig6}{a, b}).
%
%% one where the duct connecting the anterior and middle ventricles is constricted
%% (66 \% reduction in cross-sectional area);
%% one where the duct connecting the middle and posterior ventricles is constricted
%% (33 \% reduction);
%% one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
%% and the combination of these three modifications (\Crefpanels{fig:fig6}{a, b}).
%
%
In these four modified geometries, the total volume is reduced by
3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively, compared to the original
geometry. \mer{MER: HH adds a sentence on changes in tangential
  traction(\Crefpanels{fig:fig6}{d}). This is notwithstanding a
9.5\% reduction in ventral cilia force magnitude
(\Crefpanels{fig:fig6}{d}), which are the cilia that propel CSF and solute towards ROI 6.)}

Constricting the connection between the anterior and middle ventricle
reduces solute transport to the anterior ventricle (ROI 5), and the
mean concentration in ROIs 2--4 and 6 slightly increases, compared to
the original geometry (\Crefpanels{fig:fig6}{c}). \mer{We thus note
  that restricting the ...} Constricting the middle-posterior duct
mainly delays the transport into the posterior ventricle (ROI 6),
while \mer{gently} accelerating transport towards the anterior
ventricle (ROI 5). Faster transport to the anterior ventricle (ROI 5)
and slower transport to the lower middle and posterior ventricle
(ROI3, ROI 6) is also the case for the geometry with a shrunk middle
ventricle. \mer{Interestingly, this reduction in the transport to the
  posterior ventricle (ROI 6) happens in spite of a 6.6\% increase in
  the ventral cilia forces.}
%
%For the mesh where all three ventricles are modified,
When combining all three modifications,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while the mean concentration is reduced in both the anterior and the posterior ventricles (ROIs 5 and 6)s.

To summarize, these results show that the ventricular geometry and the
cross-sectional area of the ventricular ducts clearly impact solute
transport, with notable differences in transport resulting from modest
geometry modifications. These findings may explain the larger
differences in the cilia-deficient computational model results
compared to the experimental data in \Cref{fig:fig4}, \mer{as the
computational geometry is not an exact representation for the
individual zebrafish ventricles.} 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}
